---
title: "R Notebook"
output: rmarkdown::github_document
---




```{r}
library(tidyverse)
library(ggplot2)
```
```{r}
districts = read_csv("./data/districts.csv")
districts
```
There are 146 rows and 27 columns in districts.

Q2. Notice that the first row corresponds to the whole State of Tennessee. Remove this row and save the result back to districts.

```{r}
districts = districts[-1, ]
head(districts)
```

### Q3. How many districts have a proficiency rate of at least 80% for both alg_1 and eng_1?

```{r}
districts%>%filter(alg_1 >= 80 & eng_1 >= 80)
```

```{r}
count(districts%>%filter(alg_1 >= 80 & eng_1 >= 80))
```
There are 13 districts.


### Q4. How many districts have a proficiency rate less than 50% for either alg_1 or eng_1?
```{r}
districts%>%filter(alg_1 < 50  | eng_1 < 50)
```
```{r}
count(districts%>%filter(alg_1 < 50  | eng_1 < 50))
```

There are 8 districts. 


### Q5.Which district has the lowest graduation rate?
```{r}
districts%>%filter(grad == min(districts$grad, na.rm = TRUE))%>%
  select(system_name, grad)
```
Tennessee School for Blind has the lowest graduation rate. 


### Q6.  Within the Mid Cumberland region, which district has the highest ACT composite?
```{r}
districts%>%filter(region == 'Mid Cumberland')%>%
  filter(act_composite == max(act_composite, na.rm = TRUE))%>%
  select(system_name, act_composite)
```

### Q7. Create a histogram showing the distribution of graduation rates. What can you say about this distribution?
```{r}
ggplot(districts, aes(grad)) +
  geom_histogram(binwidth = 5, fill = 'blue', color = 'black')+
  scale_x_continuous(breaks=seq(0, 100, 10))
  
``` 
Most graduation rates fall between 80-100%. There is one outlier with a graduation rate 11%.

### Q8. Create a scatter plot to compare alg_1 proficiency rates to alg_2 rates. What do you notice? 
###     Facet this plot by region. Does anything stand out when you facet the plots?


```{r}
ggplot(districts) +
  geom_point(aes(alg_1, alg_2), color = 'blue') +
  scale_x_continuous(breaks=seq(0, 100, 10))
```
The scatter plot shows there is a strong correlation between alg_1 and alg_2 proficiency rates. 


```{r}
ggplot(districts) + 
  geom_point(aes(alg_1, alg_2, color = factor(region))) +
  facet_wrap((.~region), nrow = 3, ncol = 3)
```
The scatter plots show Southwest/Memphis shows the strongest positive correlation.Southeast shows the weakest correlation.
Upper Cumberland shows a non-linear trend. 


### Q9. Create a bar chart showing the total enrollment by region. 
###     Which region has the highest total enrollment? Which has the smallest?



```{r}
options(scipen = 999)
p = ggplot(districts, aes(x=region, y=enrollment, fill = region)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Dark2")

p + coord_flip() 
  
  
```

Reordering by enrollment

```{r}
options(scipen = 999)
p = ggplot(districts, aes(x = reorder(region, +enrollment), y=enrollment, fill = region)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Dark2")

p + coord_flip() 

```
upper cumberland and southwest memphis do not get in order. 
Mid Cumberland has the highest enrollment, Northwest has the lowest enrollment. 

### Q10. When creating this bar chart you may have noticed that some districts have missing enrollment values. 
###      For how many districts is this the case?

```{r}

sum(is.na(districts$enrollment))

```

```{r}
districts%>%filter(is.na(enrollment))%>%group_by(region, system_name)%>%tally()
```
### Q11. What is the mean graduation rate across all districts? 
###      What might be wrong with using just the regular mean to assess average graduation rates?

```{r}
mean(districts$grad, na.rm = TRUE)

```


### Q12. Redo the previous question but use a weighted average (weighted.mean) graduation across all districts, weighing by enrollment. 
###      How much does this change your answer? 
###      Can you explain using the data the reason for the big change from using the mean?

```{r}
weighted.mean(districts$grad, dplyr::coalesce(districts$enrollment,0), na.rm=TRUE)
```

```{r}
df <- districts %>% filter((!is.na(enrollment))&(!is.na(grad)))

```
```{r}
weighted.mean(df$grad, df$enrollment)
```

The unweighted graduation rate is 90%. 
The  enrollment weighted graduation rate is 87%.
Enrollment weighted -- schools with larger enrollments influence
the average more. 
The weighted average decreased--schools with larger enrollments have
smaller graduation rates.
Weighted-higher enrollments contribute more to the avg.

### Q13. Create a boxplot showing graduation rates per region. 
###      Does anything stand out?

```{r}
ggplot(districts)+
  geom_boxplot(aes(x=region, y=grad, color = region))+
  ggtitle('Graduation Rate by Region')+
  coord_flip()+
  theme_minimal()
```
The boxplot shows outliers for Southwest/Memphis at ~50% and MidCumberland at ~10%


### Q14. Find the weighted average of graduation rates by region using enrollment as weights. 
### Compare the results you get for the weighted average to what you see from the boxplots.


```{r}
options(digits = 3)
districts %>% 
  group_by(region)%>%
  summarise(enrollment_weighted_grad = weighted.mean(grad, dplyr::coalesce(enrollment,0), na.rm=TRUE))
  
  
  
  
```

Southwest/Memphis region has a lower graduation rate of 79.5, which is consistent with its boxplot
showing a small outlier and skewed left distribution.
Midcumberland's graduation rate is not that much lower than other regions even though its boxplot
shows a small outlier around ~12%.


### Q15. For many districts, values for alg_2 are lower than for alg_1. 
 
```{r}
districts <- districts %>% mutate(alg_diff = alg_1 - alg_2)

```


### Create a histogram showing the distribution of differences (alg_1 - alg_2).

```{r}
ggplot() + 
  geom_histogram(aes(x=alg_diff, fill="a"), alpha=.7, data=subset(districts, alg_diff <=0),  color = "blue", binwidth = 5) +
  geom_histogram(aes(x=alg_diff, fill="z"), alpha=.5, data=subset(districts, alg_diff >0),  color="black", binwidth = 5) +
  
  scale_fill_manual(name="", values=c("a" = "skyblue", "z" ="red"), labels=c("a"="Alg1 < Alg2", "z" ="Alg1 > Alg2"))+
  
  scale_x_continuous(breaks=seq(-20, 60, 10))
  

```
The histogram shows about 20 districts had a drop in algebra by about 15 points. 


### Which school had the largest drop from alg_1 to alg_2? 

```{r}
districts%>%filter(alg_diff == max(alg_diff, na.rm=TRUE))%>%select(system_name, alg_1, alg_2, alg_diff)
```
Pickett County has the largest drop from alg_1 to alg_2.


### For what percentage of schools is it true that alg_2 is larger than alg_1? 
```{r}
districts%>%summarise(pct_sch = sum(alg_diff<0, na.rm = TRUE)/nrow(districts))
```
```{r}
districts%>%summarise(pct_sch = 100*sum(na.omit(alg_diff<0))/145)
```
12.4% of schools have alg_2 greater than alg_1. 


### Is there a similar drop off for eng_2 and eng_3?

```{r}
districts <- districts %>% mutate(eng23_diff = eng_2 - eng_3)

```

```{r}
ggplot() + 
  geom_histogram(aes(x=eng23_diff, fill="a"), alpha=.9, data=subset(districts, eng23_diff <0),  color = "blue", binwidth = 5) +
  geom_histogram(aes(x=eng23_diff, fill="z"), alpha=.5, data=subset(districts, eng23_diff >0),  color="black", binwidth = 5) +
  
  scale_fill_manual(name="", values=c("a" = "skyblue", "z" ="red"), labels=c("a"="Eng2 < Eng3", "z" ="Eng2 > Eng3"))
  
  #scale_x_continuous(breaks=seq(-20, 60, 10))
  

```
The histogram shows almost all schools had drops in scores from Eng2 to Eng3. 


```{r}
districts%>%summarise(pct_sch = 100*sum(na.omit(eng23_diff>0))/145)
```
86% of schools showed decrease in score from eng2 to eng3. 


### 16. You may have noticed that a lot of rows are missing values. 
###     Which district has the largest number of missing values? 
###     What do you notice about schools that have a lot of missing values?



```{r}
districts %>% mutate(na_ct = rowSums(is.na(.)))%>%filter(na_ct == max(na_ct))
```
West Tenn School for Deaf has the most missing values


```{r}
districts %>% mutate(na_ct = rowSums(is.na(.)))%>%arrange(desc(na_ct))
```

I cannot see anything that stands out...yet. 

###Q17. Find the correlation between graduation rate and all other variables.
###     Create a horizontal bar chart showing these correlations. 
###     Make sure that your plot is ordered by correlation values. 
###     What do you notice from these correlations?



```{r}
districts %>% 
  select(where(is.numeric)) %>% 
  cor(. , districts$grad, use = 'complete.obs')
```


```{r}
cor_dt <- districts %>% 
  select(where(is.numeric)) %>% 
  cor(. , districts$grad, use = 'complete.obs') %>% 
  as_tibble(cor_dt, rownames = "variable") %>% 
  rename(correlation = V1)%>%
  filter(variable != "grad")


cor_dt
```

```{r}
ggplot(data = cor_dt, aes(y=reorder(variable, -correlation), x=correlation)) +
  geom_bar(stat = 'identity')+
  theme_bw()



```


```{r}
ggplot(data = cor_dt, aes(x=correlation, y=reorder(variable, -correlation), fill=correlation)) +
  geom_bar(stat = 'identity')+
  scale_fill_gradient(low='firebrick', high='skyblue', space='Lab')




```

Dropout and suspension rates are the most negatively correlated with graduation rate(surprise!)
Scores in English, Algebra and Sciences are the most positively correlated with graduation rates.


###Q18. Create a scatterplot for grad vs. suspended. 
###     Does what you see make sense given your answer from the previous part?


```{r}
ggplot(districts) +
  geom_point(aes(suspended, grad),  na.rm = TRUE)+
  xlim(0,50)+
  ylim(50,100)
  
```
The scatterplot shows that lower suspension rates are associated with higher graduation rates. 





